---
title: "Random Forest for Biometric Data"
author: "Brandon Hansen"
date: "7/31/2019"
output: html_document
---

Load the data into R to be used with the generalized linear model.
Divide the data into a training and testing sets.

```{r}
library(caret)
training <- na.exclude(read.csv("pml-training.csv"))
training$classe <- as.factor(training$classe)
testing <- read.csv("pml-testing.csv")
```

Specify the control for the cross validation.
In this case I'm using 20 fold cross validation.
```{r}
control <- trainControl(method="cv", number=20)
```

Model was built using random forest as the machine learning algorithm.
Feature selection was determined from seperate experiments with differing feature combinations.
Feature selection was also determined by feature presence in test dataset.
```{r}
ml <- train(classe ~ 
                roll_belt + pitch_belt + yaw_belt
              + accel_belt_x + accel_belt_y + accel_belt_z
              #+ gyros_belt_x + gyros_belt_y + gyros_belt_z
              + magnet_belt_x + magnet_belt_y + magnet_belt_z
            
              #+ roll_arm + pitch_arm + yaw_arm 
              #+ accel_arm_x + accel_arm_y + accel_arm_z
              #+ gyros_arm_x + gyros_arm_y + gyros_arm_z
              #+ magnet_arm_x + magnet_arm_y + magnet_arm_z
            
              + roll_dumbbell + pitch_dumbbell + yaw_dumbbell
              + accel_dumbbell_x + accel_dumbbell_y + accel_dumbbell_z
              #+ gyros_dumbbell_x + gyros_dumbbell_y + gyros_dumbbell_z
              + magnet_dumbbell_x + magnet_dumbbell_y + magnet_dumbbell_z
            
              + roll_forearm + pitch_forearm + yaw_forearm
              + accel_forearm_x + accel_forearm_y + accel_forearm_z
              #+ gyros_forearm_x + gyros_forearm_y + gyros_forearm_z
              + magnet_forearm_x + magnet_forearm_y + magnet_forearm_z, 
              model = "rf", data = training,
              trControl = control)

print(ml)
```
Cross validation was not only used to test the models accuracy but to also select the best model out the individual ones generated.

Based on results from the 20 fold cross validation, the expected out of sample error should be similar to the error found above.

Generation of predictions for test dataset.
```{r}
predict(ml, testing)
```

